---
title: "Data formatting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data formatting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup, warning=FALSE, message=FALSE}
library(ForestPlotR)
library(dplyr)
```

This vignette describes how mapped forest stand datasets should be formatted to
ensure full compatibility with ForestPlotR and introduces some data checking
functions to assist with the formatting process.

## Data file types

The ForestPlotR package expects raw mapped forest stand data to consist of at 
least two distinct files:

1. Mapping data - coordinate locations of individual trees within stands

2. Tree census data - measurements of individual trees taken during repeated
censuses of the stands

### Mapping data

The mapping dataset should be a data frame where each row represents a unique
tree and contains its identifying and location information. The built-in
dataset `mapping` is a cleaned and correctly formatted example:

```{r}
head(mapping)
```

To ensure full compatibility with ForestPlotR functions, the mapping dataset
should contain the columns: `tree_id`, `stand_id`, `tag`, `species`, `x_coord`,
and `y_coord`. Additional columns can also be included without disrupting
function performance and column order is not important. For details on what each
required column represents, use `?mapping`. If only a few functions from
ForestPlotR will be  applied, some columns in `mapping` may not be required (e.g.
the column `tag` is only required for `stand_map()`) so please read
individual function documentation before spending time generating any missing
columns.

### Tree census data

The tree census dataset should be a data frame where each row represents the 
measurement data for a single tree during a specific census of the stand. This
means each tree will appear on *x* rows where *x* is the number of censuses in 
which that tree was measured. The required columns are: `tree_id`, `stand_id`,
`species`, `year` and `dbh`. For details on what each required column
represents, use `?tree`. The built-in dataset `tree` is a cleaned and
correctly formatted example:

```{r}
head(tree)
```

Additional unrequired columns can also be included in the tree census dataset
without disrupting function performance (e.g. `tag` in `tree`) and column order
is not important.

## Data checking

To obtain accurate and useful neighborhood descriptions it is important that
the mapping and tree census datasets are cleaned of any unusual or missing data.
The ForestPlotR package contains functions to assist with the cleaning process
but these functions only highlight cases of missing data rather than 
automatically removing the associated observations. This is because not all
cases of missing data prevent all types of analysis and some missing data can
be inferred from data collection records (e.g. missing year of measurement in
tree census data).

### `mapping_check()`: Check mapping datasets

The built-in dataset `messy_mapping` contains examples of common data errors in
mapping datasets. For instance, there are 10 tree ids that are connected to more
than one mapping record:

```{r}
messy_mapping %>%
  group_by(tree_id) %>%
  summarize(count = n()) %>%
  filter(count > 1)
```

The `mapping_check()` function checks a mapping dataset for a variety of common
errors and returns a list containing two elements. The first element 
(`$problem_trees`) is a data frame containing the rows of the input mapping
dataset that contain issues, with an additional final column describing the
issue. The arguments "max_x" and "max_y" must be provided so the function can
check for x and y coordinates beyond the stand boundary.

```{r}
map_issues <- mapping_check(messy_mapping, max_x = 100, max_y = 100)
head(map_issues$problem_trees)
```

The second element (`$issue_summary`) is a data frame summarizing the number and
percentage of trees in the mapping dataset that have one or more problems and
each specific type of problem.

```{r}
head(map_issues$issue_summary)
```

### `tree_check()`: Check tree census datasets

The built-in dataset `messy_tree` contains examples of common data errors in
tree census datasets. For instance, there are 10 tree measurement records that
have no dbh information:

```{r}
messy_tree %>%
  filter(is.na(dbh))
```

The `tree_check()` function checks a tree census dataset for a variety of common
errors. A mapping dataset needs to be supplied to the function so that trees in
the tree census data that have no associated mapping can be identified. 
`tree_check()` returns a list containing two elements. The first element 
(`$problem_trees`) is a data frame containing the tree ids that were flagged as
having a data issue and summarizes the issue.

```{r}
tree_issues <- tree_check(tree_data = messy_tree, map_data = mapping)
head(tree_issues$problem_trees)
```

The second element (`$issue_summary`) is a data frame summarizing the number and
percentage of trees in the tree census dataset that have one or more problems
and each specific type of problem. Note that a tree id will be flagged if just
one of its measurement records contains an issue, so many flagged trees are
likely to be usable for most analyses. 

```{r}
head(tree_issues$issue_summary)
```
