---
title: "Describing neighborhoods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Describing neighborhoods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup, warning=FALSE, message=FALSE}
library(ForestPlot)
library(dplyr)
```

This vignette describes a series of functions in the ForestPlot package that 
provide quantitative descriptions of local neighborhoods in mapped forest
stands. The neighborhood of a specific location in a mapped stand is the 
collection of trees growing within a specified distance of that location. 

Of these functions, `neighborhoods()` constructs the list of all trees in each
neighborhood. `neighborhood_summary()` takes the output of `neighborhoods()` as
input and calculates tree species richness and species-specific densities for
each neighborhood. `site_by_species()` also takes the output of
`neighborhoods()` as input but creates a site-by-species matrix, where each site
is a different neighborhood, that can be used to calculate diversity metrics.
`tree_utm()` takes a mapping object as input and calculates UTM coordinates for
each tree, thereby allowing trees to be linked to fine-scale spatial data such
as topography.

## `neighborhoods()`: Construct neighborhoods

The `neighborhoods()` function uses a mapping dataset to determine, for each
tree, all the trees that are within a certain distance (`radius`). The argument
`stand` can be used to indicate that neighborhoods should be constructed only
for the trees of one or more of the stands included in `mapping`. If `stand` is
not specified, neighborhoods will be constructed for all trees in `mapping`.

The output is a data frame where each row contains information on a focal tree
and one other tree growing in the focal tree's neighborhood. Therefore, the
information for each focal tree will appear on *x* lines, where *x* is the
number of trees in that focal tree's neighborhood. This structure is shown
below by isolating a single focal tree in the output of applying 
`neighborhoods()` to the built-in `mapping` dataset:

```{r}
nbhds <- neighborhoods(mapping, stand = "AB08", radius = 10)
nbhds %>%
  filter(tree_id == "AB08000100022")
```

The `neighborhoods()` function can also be used to construct neighborhoods for
specific coordinates within the mapped stand instead of constructing a
neighborhood centered on each tree in the dataset. This is achieved by
providing a data frame of coordinates under the optional argument `coords` (see
`neighborhoods()` documentation for details on how this data frame should be
structured).

```{r}
# Create a data frame of coordinates
locations <- data.frame(
  loc_id = paste("A", 1:81, sep = ""),
  x_coord = rep(seq(10, 90, 10), times = 9),
  y_coord = rep(seq(10, 90, 10), each = 9))

# Construct neighborhood for each coordinate
coord_nbhds <- neighborhoods(mapping, stands = "AB08", radius = 10,
                             coords = locations)
head(coord_nbhds)
```

## `neighborhood_summary()`: Summarize neighborhoods

The `neighborhood_summary()` function takes output of the `neighborhoods()`
function and calculates species richness, overall tree density, and densities of
each tree species for each neighborhood. The `densities` argument is used to
specify how densities should be calculated.

### `neighborhood_summary(densities = "raw")`

Overall tree density and species-specific densities will be presented in units
of $m^{2}ha^{-1}$, with the area occupied by each tree calculated as the area
of the trunk at breast height i.e.
$$\pi(\frac{dbh}{2})^{2}$$
```{r}
nbhd_summ <- neighborhood_summary(nbhds, id_column = "tree_id", radius = 10,
                                  densities = "raw")
nbhd_summ %>%
  select(tree_id, sps_richness, ABAM_density, TSHE_density, all_density)
```

### `neighborhood_summary(densities = "proportional")`

Overall tree density is presented in units of $m^{2}ha^{-1}$ but species-specific densities represent the proportion of overall tree density that each species
accounts for.

```{r}
nbhd_summ <- neighborhood_summary(nbhds, id_column = "tree_id", radius = 10,
                                  densities = "proportional")
nbhd_summ %>%
  select(tree_id, sps_richness, ABAM_density, TSHE_density, all_density)
```

### `neighborhood_summary(densities = "angular")`

Overall tree density and species-specific densities calculated as the sum of 
angles occupied by trees, where the angle occupied by a single tree is:
$$arctan(\frac{dbh}{distance})$$ In the above, `dbh` is the dbh of the
neighbor tree and `distance` is the distance from the neighborhood center to
the neighbor tree.

```{r}
nbhd_summ <- neighborhood_summary(nbhds, id_column = "tree_id", radius = 10,
                                  densities = "angular")
nbhd_summ %>%
  select(tree_id, sps_richness, ABAM_angle_sum, TSHE_angle_sum, all_angle_sum)
```

## `site_by_species()`: Create site-by-species matrices

The `site_by_species()` function takes output of the `neighborhoods()` function
and creates a site-by-species matrix where each site is a neighborhood. The
resulting matrix can be used as input for functions in other packages (e.g.
picante, vegan) that calculate various diversity metrics from site-by-species
matrices.

The default is to present presence/absence data in the matrix:

```{r}
ss_mat <- site_by_species(nbhds, id_column = "tree_id")
head(ss_mat)
```

But if `abundance = T` is specified, matrix element values will represent the
number of trees of the associated species in the neighborhood:

```{r}
ss_mat_abund <- site_by_species(nbhds, id_column = "tree_id", abundance = T)
head(ss_mat_abund)
```

## `tree_utm()`: Calculate tree UTM coordinates

The `tree_utm()` function takes a mapping object as input and calculates UTM
coordinates for each tree. The output spatial object can be used to extract
further data for each tree from data sources with very high spatial resolution
e.g. topography.

The required arguments for `tree_utm()` are: geographic coordinates of the 
stands (`stand_locs`), the ESPG code of the coordinate reference system of 
`stand_locs` (`original_crs`), and the ESPG code of the UTM zone in which the 
stands exist (`utm_crs`).

The built-in `stand_locations` dataset shows the expected structure of 
`stand_locs` - see `?stand_locations` for details:

```{r}
head(stand_locations)
```

The output of `tree_utm()` looks like the following:

```{r}
mapping_utm <- tree_utm(mapping, stand_locations, 4326, 32610)
head(mapping_utm)
```

