---
title: "Modeling tree growth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling tree growth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
library(dplyr)
library(glmnet)
```

```{r setup}
library(ForestPlot)
```

This vignette demonstrates how the neighborhood description and growth
calculation functions of ForestPlot can be used to setup a neighborhood model
of tree growth.

## Formating data for a growth model

To build a neighborhood model of tree growth, we first need to describe tree
neighborhoods. The ForestPlot functions `neighborhoods()` and
`neighborhood_summary()` make this very simple.

```{r}
# Construct neighborhoods for all trees in all stands
nbhds <- neighborhoods(mapping, radius = 10)

# Describe neighborhoods using angular species-specific densities
nbhd_summ <- neighborhood_summary(nbhds, id_column = "tree_id", radius = 10,
                                  densities = "angular")

# Combine neighborhoods with their summaries
nbhds <- nbhds %>%
  left_join(nbhd_summ, by = "tree_id")
```

We have now summarized the neighborhoods of all trees, but our summaries are 
inaccurate for trees whose neighborhood overlaps a stand boundary i.e. we did
not sample their entire neighborhood. Rather than estimating the missing
portions of these trees' neighborhoods, we will exclude them from the model.

```{r}
# Define neighborhood radius
nb_rad <- 10

# Remove trees whose neighborhood overlaps a stand boundary
nbhds <- nbhds %>%
  filter(x_coord >= nb_rad & x_coord <= 100 - nb_rad &
           y_coord >= nb_rad & y_coord <= 100 - nb_rad)
```

Next we calculate annual growth for our trees. We will also remove any trees
that were only measured once (because we cannot calculate a growth rate for
them) or had a nonsensical negative growth rate.

```{r}
# Calculate annual growth for all trees
growth <- growth_summary(tree)

# Remove trees that were only measured once and/or had negative growth
growth <- growth %>%
  filter(growth$first_record != growth$last_record &
           annual_growth >= 0)
```

Now we can add the growth rates to the neighborhoods data. For our model, we
will use `size_corr_growth` because this is the closest to being normally
distributed. When combining we use `inner_join()` because not all trees in
`nbhds` have growth data e.g. trees measured only one time.

```{r}
nbhds <- nbhds %>%
  inner_join(growth %>% select(tree_id, size_corr_growth),
             by = "tree_id")
```

We can also include additional covariates in our model such as abiotic data.
The code below adds abiotic data to `nbhds` from the built-in dataset 
`stand_abiotic`.

```{r}
abio <- stand_abiotic
nbhds <- nbhds %>%
  left_join(abio, by = "stand_id")
```

Finally, we can drop some of the columns from `nbhds` that we don't need for
our model.

```{r}
nbhds <- nbhds %>%
  select(-c(stand_id, dbh, abh, x_coord, y_coord, id_comp, abh_comp))
```

Now we have a dataset containing all the variables we will use for this
modeling demonstration.

## `growth_model()`: Create a neighborhood tree growth model

When modeling tree growth, it is common practice to create a separate model
for each focal tree species because the drivers of growth are likely to differ
among species. The code below isolates a single tree species and then removes
the species variable because it is no longer useful.

```{r}
one_species <- nbhds %>%
  filter(species == "ABAM") %>%
  select(-species)
```

The `growth_model()` function also offers test set validation, so next we will
divide our dataset into a training and test set.

```{r}
# Create vector of tree ids
all_tree_ids <- unique(one_species$tree_id)

# Assign 80% of trees to training
training_trees <- sample(all_tree_ids, length(all_tree_ids) * 0.8)

# Define training and test
training <- one_species %>%
  filter(tree_id %in% training_trees)
test <- setdiff(one_species, training)
```

Now we can call the `growth_model()` function, which fits a linear regularized
regression model of tree growth. In the most basic use case, this function
requires only a training set and the name of the dependent variable, which 
should be a column in the training set. We set the seed because the model fitting process is stochastic.

```{r}
set.seed(1000)
basic_growth_mod <- growth_model(training, outcome_var = "size_corr_growth")
```

The output of the above code is a list containing 4 elements. The first element
`mod` is the model object. The second element `obs_pred` is a data frame of
observed and model predicted growth for each tree. The third element is the
R^2 value of the model. The fourth element is the estimated coefficients.

```{r}
# R^2 value
basic_growth_mod$R_squared

# Observations and model predictions
basic_growth_mod$obs_pred
```

The evaluate the robustness of the stochastic model fitting process there is an
optional argument (`iterations`) that specifies how many times the model should
be fit. When this argument is specified, the model object, r-squared value and
table of observations and predictions refer to the model that had the lowest
mean square error. However, the `mod_coef` element will contain the fitted
coefficients for all fitted models.

```{r}
set.seed(1000)
iter_growth_mod <- growth_model(training, outcome_var = "size_corr_growth",
                                iterations = 10)
iter_growth_mod$R_squared
```

The `growth_model()` function can also handle rare competitor tree species. The
optional argument `rare_comps` specifies the minimum number of interactions a
competitor species must be a part of to be considered a common competitor. Any
species with fewer interactions will be grouped together under the new species
identity of `"RARE"`. If species-specific densities are included as covariates
in the training data, a `"RARE_density"` variable can also be created by
indicating the suffix attached to density variables under the argument 
`density_suffix`.

```{r}
set.seed(1000)
rare_growth_mod <- growth_model(training, outcome_var = "size_corr_growth",
                                rare_comps = 100, density_suffix = "_angle_sum")

# Check for RARE in coefficients
names(rare_growth_mod$mod_coef)
```

Finally, the `growth_model()` function can evaluate the ability of the fitted
model to predict the growth of trees in a test set i.e. trees not used in the
model fitting process. To do this, a test set must be provided under the 
optional argument `test`. Predictive ability is quantified with an R^2 value
that constitutes a new element in the output list.

```{r}
set.seed(1000)
test_growth_mod <- growth_model(training, outcome_var = "size_corr_growth",
                                test = test)

# Ability to predict test data
test_growth_mod$test_R_squared
```
