---
title: "Selecting neighborhood size"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Selecting neighborhood size}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup, warning=FALSE, message=FALSE}
library(ForestPlotR)
library(dplyr)
```

This vignette demonstrates how to select an appropriate neighborhood size for
modeling tree growth. The appropriate neighborhood size differs between forest
ecosystems, forest successional stages, and tree species, so it is strongly
recommended that a sensitivity analysis be conducted for neighborhood size
before drawing conclusions.

## Using `select_nbhd_size()`

The `select_nbhd_size()` function fits tree growth models using a variety of
neighborhood sizes to calculate the explanatory neighborhood metrics and
quantifies the fit of each model to the training data. Comparing the fit of 
models based on different neighborhood sizes then permits an informed decision
on the neighborhood size to be used in the final analysis.

This process involves the implementation of many ForestPlotR functions and 
therefore `select_nbhd_size()` has many required arguments. First we must 
provide the neighborhood sizes to try (`radii`) and the tree species for which
we want to try them (`focal_sps`; independent assessments of neighborhood size
are made for each species). In addition, we must provide the tree location data
required for identifying neighborhoods (`map_data`), the tree remeasurement
data needed to calculate growth rates (`growth_data`), and optionally, any
stand-level abiotic variables we want to include as covariates in the models
(`abiotic_data`).

We also need to indicate what method `neighborhood_summary()` should use to
calculate species-specific tree densities, the maximum x and y coordinates so
that trees with neighborhoods overlapping the stand boundary can be excluded
(`max_x` and `max_y`), and (optionally) the number of interactions a neighbor
species must be involved in to be treated independently instead of included in
the "RARE" group (`rare_sps`; defaults to 100 interactions if not provided).

The following code shows an example of applying `select_nbhd_size()` to the
species "PSME" and "TSHE" in the built-in mapping data, trying neighborhood
sizes of 3, 6, 9, 12, and 15 m in radius:

```{r}
size_compare <- select_nbhd_size(radii = seq(3, 15, 3), map_data = mapping,
                                 growth_data = tree,
                                 abiotic_data = stand_abiotic,
                                 focal_sps = c("PSME", "TSHE"),
                                 dens_type = "proportional",
                                 max_x = 100, max_y = 100)
```

The output of `select_nbhd_size()` is a list. The first element is a table of
the mean square errors of the models based on different neighborhood sizes
applied to each of the focal species:

```{r}
size_compare$mse_vals
```

The remaining elements are plots of mean square error vs. neighborhood size
for each of the focal species:

```{r}
size_compare$PSME_plot
```

```{r}
size_compare$TSHE_plot
```

To select the most appropriate neighborhood size using these plots, we need to
consider two competing priorities. First, we of course want our model to fit
the training data well so we want to pick a neighborhood size that results in
a low mean square error. However, we also want to include as many focal
trees as we can because this gives us greater confidence that our findings will
be applicable beyond the sampled stands.

As a larger neighborhood size results in lower mean square error but more focal
trees excluded due to their neighborhood overlapping the stand boundary, this
decision presents a trade-off. We therefore recommend choosing the neighborhood
size that, based on visual observation, lies at the “elbow” of these plots. In
the example shown here, we would select 6m as the neighborhood radius for both
PSME and TSHE. We would therefore provide 6m as the neighborhood radius to the
`neighborhoods()` function when preparing the data for our final model.
