---
title: "ForestPlot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ForestPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The ForestPlot R package is designed to facilitate the cleaning, analysis and
visualization of mapped forest plot data.

In this vignette we: (1) describe the intended use cases of ForestPlot, (2)
explain how datasets need to be formatted to be compatible with ForestPlot
functions, and (3) use the built-in datasets to demonstrate each included
function.

## Intended uses of ForestPlot

In mapped forest plots, the precise locations of individual trees are recorded.
This means that for any location within the plot, we can describe the local tree
community in a spatially explicit way by identifying the size, proximity and 
species identity of nearby trees. These descriptions are termed neighborhoods
and can be used as covariates in models of forest processes such as tree
growth/mortality or nutrient cycling. The main goal of ForestPlot is to provide
functions that can efficiently quantify and explore these neighborhoods. 

Specific questions ForestPlot could be used to answer:

* How are growth/survival rates of trees influenced by neighboring trees?
* What is the relationship between soil microbial community composition and
tree neighborhood?
* Where in this mapped stand is tree density the lowest?

## Data formatting requirements





```{r setup}
library(ForestPlot)
```
