---
title: "Visualizing mapped stands"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing mapped stands}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.align = "center")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup, warning=FALSE, message=FALSE}
library(ForestPlotR)
library(dplyr)
```

This vignette describes the ways in which mapped forest stand data can be 
visualized using the ForestPlotR package.

The `size_dist()` function can be used to produce and compare size-class
distribution figures. `contour_plot()` uses measurements of any quantitative
variable associated with coordinates to fit a contour map of that variable over
a mapped stand. `stand_map()` produces a map of all trees in a mapped stand that
can be used to navigate in the field.

## `size_dist()`: Explore size-class distributions

The `size_dist()` function uses a tree census dataset to produce a size-class
distribution of a specified stand. The most recent measurement of each tree in
the census dataset is used to assign it to a size class. Below is an example
of applying `size_dist()` to a single stand in the built-in `tree` dataset:

```{r}
size_dist(size_data = tree, stands = "AB08")
```

The default width of size-classes in the resulting plot is 10 dbh units, but
can be changed:

```{r}
size_dist(size_data = tree, stands = "AB08", bin_size = 5)
```
 
It is also possible to produce size-class distributions for multiple stands in
order to compare them. To ensure clarity of individual distributions it is 
recommended that fewer than 10 stands be compared at once.

```{r}
size_dist(size_data = tree, stands = c("AM16", "AO03", "AR07"))
```

## `contour_plot()`: Interpolate contours across stands

We often measure variables at known coordinates within mapped forest stands. If 
measurements are taken at enough locations, it may be desirable to interpolate
the measured variable across the entire stand. This is the aim of the
`contour_plot()` function.

With mapped forest stand data it is also possible to derive measurements at 
coordinate locations by calculating features of neighborhoods centered at those
coordinates. As an example, here we will calculate tree density across a 
coordinate grid and create a contour plot of tree density.

First we create a grid of coordinate locations:

```{r}
locations <- data.frame(
  loc_id = paste("A", 1:441, sep = ""),
  x_coord = rep(seq(0, 100, 5), times = 21),
  y_coord = rep(seq(0, 100, 5), each = 21))
```

Then we calculate tree density at each of these locations using ForestPlotR
functions:

```{r}
nbhds <- neighborhoods(mapping, stands = "AB08", radius = 10,
                       coords = locations)
nbhd_summ <- neighborhood_summary(neighbors = nbhds, id_column = "loc_id",
                                  radius = 10, densities = "angular")
```

Next we join the coordinate locations with the calculated densities:

```{r}
grid_data <- locations %>%
  left_join(nbhd_summ %>%
              select(loc_id, all_angle_sum),
            by = "loc_id") %>%
  select(-loc_id)
```

Now we can make the contour plot:

```{r}
contour_plot(grid_data, "all_angle_sum", rad = 10, max_x = 100, max_y = 100)
```

In the above plot, density appears to decrease toward the stand edges. This is
because the locations close to the stand boundary did not have their entire
neighborhood sampled. To improve the accuracy of values for these edge 
neighborhoods, we can use the `edge_correction` argument of
`neighborhood_summary()`, which multiplies density values according to the
proportion of the neighborhood area that lies beyond the stand boundary:

```{r}
nbhd_summ <- neighborhood_summary(nbhds, id_column = "loc_id", radius = 10,
                                  densities = "angular", edge_correction = T,
                                  x_limit = 100, y_limit = 100)
grid_data <- locations %>%
  left_join(nbhd_summ %>%
              select(loc_id, all_angle_sum),
            by = "loc_id") %>%
  select(-loc_id)
contour_plot(grid_data, "all_angle_sum")
```

## `stand_map()`: Create maps of forest stands

The `stand_map()` function creates a map of a forest stand with each tree
represented by a point sized relative to its `dbh` and labeled with its `tag`
number. The x and y coordinate limits must also be provided and can be used
to create a map of only a portion of the stand:

```{r, fig.height = 6, fig.width = 6}
# Isolate mapping and tree data for one stand
one_stand_map <- mapping %>%
  filter(stand_id == "AB08")
one_stand_tree <- tree %>%
  filter(stand_id == "AB08")

# Create stand map
stand_map(one_stand_map, one_stand_tree, c(50, 100), c(50, 100), "tag")
```
