---
title: "Visualizing mapped plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing mapped plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.align = "center")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup, warning=FALSE, message=FALSE}
library(ForestPlot)
library(dplyr)
```

This vignette describes the ways in which mapped forest plot data can be 
visualized using the ForestPlot package.

The `size_dist()` function can be used to produce and compare size-class
distribution figures. `contour_plot()` uses measurements of any quantitative
variable associated with coordinates to fit a contour map of that variable over
a mapped plot `stand_map()` produces a map of all trees in a mapped plot that
can be used to navigate in the field.

## `size_dist()`: Explore size-class distributions

The `size_dist()` function uses a tree census dataset to produce a size-class
distribution of a specified stand. The most recent measurement of each tree in
the census dataset is used to assign it to a size class. Below is an example
of applying `size_dist()` to a single stand in the built-in `tree` dataset:

```{r}
size_dist(size_data = tree, stands = "AB08")
```

The default width of size-classes in the resulting plot is 10 dbh units, but
can be changed:

```{r}
size_dist(size_data = tree, stands = "AB08", bin_size = 5)
```
 
It is also possible to produce size-class distributions for multiple plots in
order to compare them. To ensure clarity of individual distributions it is 
recommended that fewer than 10 plots be compared at once.

```{r}
size_dist(size_data = tree, stands = c("AM16", "AO03", "AR07"))
```

## `contour_plot()`: Interpolate contours across plots

We often measure variables at known coordinates within mapped forest plots. If 
measurements are taken at enough locations, it may be desirable to interpolate
the measured variable across the entire plot. This is the aim of the
`contour_plot()` function.

With mapped forest plot data it is also possible to derive measurements at 
coordinate locations by calculating features of neighborhoods centered at those
coordinates. As an example, here we will calculate tree density across a 
coordinate grid and create a contour plot of tree density.

First we create a grid of coordinate locations:

```{r}
locations <- data.frame(
  loc_id = paste("A", 1:441, sep = ""),
  x_coord = rep(seq(0, 100, 5), times = 21),
  y_coord = rep(seq(0, 100, 5), each = 21))
```

Then we calculate tree density at each of these locations using ForestPlot
functions:

```{r}
nbhds <- neighborhoods(mapping, stands = "AB08", radius = 10,
                       coords = locations)
nbhd_summ <- neighborhood_summary(nbhds, "loc_id", 10, "angular")
```

Next we join the coordinate locations with the calculated densities:

```{r}
grid_data <- locations %>%
  left_join(nbhd_summ %>%
              select(loc_id, all_angle_sum),
            by = "loc_id") %>%
  select(-loc_id)
```

Now we can make the contour plot:

```{r}
contour_plot(grid_data, "all_angle_sum", rad = 10, max_x = 100, max_y = 100)
```

In the above plot, density appears to decrease toward the plot edges. This is
because the locations close to the plot boundary did not have their entire
neighborhood sampled. `contour_plot()` offers a method of estimating the missing
regions of neighborhoods that overlap the plot boundary, specified by the 
`edge_handling` argument:

```{r}
contour_plot(grid_data, "all_angle_sum", edge_handling = "multiply",
             rad = 10, max_x = 100, max_y = 100)
```

## `stand_map()`: Create maps of forest plots

The `stand_map()` function creates a map of a forest plot with each tree
represented by a point sized relative to its `dbh` and labeled with its `tag`
number. The x and y coordinate limits must also be provided and can be used
to create a map of only a portion of the plot:

```{r, fig.height = 6, fig.width = 6}
# Isolate mapping and tree data for one stand
one_stand_map <- mapping %>%
  filter(stand_id == "AB08")
one_stand_tree <- tree %>%
  filter(stand_id == "AB08")

# Create stand map
stand_map(one_stand_map, one_stand_tree, c(50, 100), c(50, 100))
```
